---
title: "Review of BBMP Finances, Part 1 Trade License"
author: "Vijayvithal"
date: "April 28, 2016"
#output: ioslides_presentation
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(digits=2,scipen = 9)
library(dplyr)
library(leaflet)
library(rgdal)
lakh<-100000
crore<-10000000
```



```{r cache=TRUE}
create_tl_df<-function(){
  frame<-data.frame()
  csvfiles<- list.files(path="ziploc")
  csvdf<-data.frame()
 for (csvfile in csvfiles) {
   #print(csvfile)
   csvfile1<-paste0("ziploc","\\",csvfile)
   df<-read.csv(csvfile1,header = T,stringsAsFactors = F,colClasses = c("integer", "character", "character", "integer", "character", "character", "character", "character", "integer", "character", "character", "character", "character", "numeric")  )
   df<-mutate(df,month=csvfile)
   csvdf<-rbind(csvdf,df)
 }
  
   frame<-rbind(frame,csvdf)
   return (frame)
}
#zipfiles<-list.files(path="tabula",pattern="tabula-*",full.names = T)
df_trade_license<-data.frame()
if(!file.exists("trade_license.csv")){
 
df_trade_license<-create_tl_df()
write.csv(df_trade_license,"trade_license.csv",)
} else   {
 df_trade_license <- read.csv("trade_license.csv",header = T,stringsAsFactors = F )
}

```

```{r }
library(lubridate)
library(ggplot2)
library(xtable)
#Data Cleanup
df_trade_license$month<- gsub(".*tabula-April.*.csv","April-2015",df_trade_license$month)
df_trade_license$month<- gsub(".*tabula-Aug.*.csv","August-2015",df_trade_license$month)
df_trade_license$month<- gsub(".*tabula-Decembe.*.csv","December-2015",df_trade_license$month)
df_trade_license$month<- gsub(".*tabula-july.*.csv","July-2015",df_trade_license$month)
df_trade_license$month<- gsub(".*tabula-June.*.csv","June-2015",df_trade_license$month)
df_trade_license$month<- gsub(".*tabula-May.*.csv","May-2015",df_trade_license$month)
df_trade_license$month<- gsub(".*tabula-Nov.*.csv","November-2015",df_trade_license$month)
df_trade_license$month<- gsub(".*tabula-Octo.*.csv","October-2015",df_trade_license$month)
df_trade_license$month<- gsub(".*tabula-Sep.*.csv","September-2015",df_trade_license$month)
df_trade_license$month<- gsub(".*tabula-January.*.csv","January-2016",df_trade_license$month)
df_trade_license$month<- gsub(".*tabula-Feb-.*.csv","February-2015",df_trade_license$month)
df_trade_license$month<- gsub(".*tabula-Feb.*.csv","February-2016",df_trade_license$month)
df_trade_license$month<- gsub(".*sam_March.*.csv","March-2015",df_trade_license$month)
df_trade_license$month<- mdy(gsub("-"," 1 ",df_trade_license$month))
df_trade_license$Trade.Minor.Name<-gsub("CONSU PTION","CONSUMPTION",df_trade_license$Trade.Minor.Name )

df_trade_license<-rename(df_trade_license,WARD_NO=Ward.ID)
```
## BBMP Finances, Introduction

* For 2016-17 BBMP has put a revenue target of an additional 980 Crores from Property Tax.
* Examining the revenue potential across various segments we find that BBMP can not only meet by exceed by a large margin the target amount of 980 Crores without raising the tax rate by a single paisa. For e.g.

1. Betterment charges worth 390 Crores have not been collected in Mahadevapura Zone alone.
2. Property Taxes from 110 villages of Bellundur were not collected for two years post their merging with BBMP.
3. Sources with BBMP claim that penalties on unauthorised hoardings can raise 2,000 Crores
4. BBMP's own estimate show that if Trade license fees were collected from all traders then this year around 300 Crores can be raised instead of 18 Crores last years. If the arrears for the years for which trade license fees was not collected is recovered, the final sum will be much higher.


Over the next few month we will analyse and report the leakage in different heads of BBMP Accounts using either data sourced from BBMP or from other third party sources.

*In this report we cover the Trade license data over a 13 months period from Jan 2015 to Feb 2015*


* As per section 41a and 41b(xiii) of the RTI act all these data-sets should be proactively and regularly put in public domain.
  + We demand that the government should fulfill its obligations as per the RTI act and put these data-sets online in a format compliant with the national data sharing and accessibility policy 2012.
  + This will enable exploration detection and reporting of revenue leakage by the citizens of Bangalore

```
Q: Is the government interested in  plugging the revenue leakage.
```


## Trade License, Overview.

```{r}
ggplot(df_trade_license,aes(month,fill=Zone.Name))+geom_bar()+facet_wrap(~Zone.Name ,ncol=2)+ylab("License Application count")
```

* BBMP collects annual trade license fees from some 55,000 traders.
* By BBMP's own admission it needs to collect fees from some additional 4 lakh traders.
* An Analysis of Trade license data from 2015 shows that BBMP has either failed to collect or to report the collected trade license fees from traders in Mahadevapura and Yelhanka Zone

## We demand that 

> BBMP conduct an audit of the trade license fee collection in Mahadevapura and Yelhenka zone and put the correct figures in the public domain.


```{r, message=FALSE, warning=FALSE, include=FALSE}
revenue<-df_trade_license%>%group_by(WARD_NO)%>%summarise(Amount=sum(Amount.Paid)/lakh)
shp<-readOGR("../Data-Driven-Policy/book/bangalore/bangalore/gis/bbmpwards","bbmpwards")
shp@data<-left_join(shp@data,revenue)
shp@data<-shp@data%>%mutate(popup=paste0("Ward=",WARD_NAME,"<br/>Amount=",Amount))
```
```{r}
pal<-colorBin(rev(topo.colors(10)),shp$Amount,na.color="#FF0CC0")
leaflet(shp)%>%
addTiles(urlTemplate = "http://{s}.basemaps.cartocdn.com/light_all/{z}/{x}/{y}.png") %>%
addPolygons(stroke=T,fillColor =pal(shp$Amount), popup=shp$popup)  %>%
#addCircles(lng=shp$LON,lat=shp$LAT,radius=shp$NumPG)%>%
addLegend(position = 'topright', pal=pal,values=shp$Amount, opacity = 0.4,       title = 'Trade license fees(in lakhs)/Ward')
```
```

Q: Why is Shantala Nagar able to collect 1 Crore+ while Market areas like Malleshwaram (4 lakh), Shivaji Nagar(5 lakh),Chickpet(11 lakh), K.R Market(6 lakhs) collect less than a 10th?
```
## We Demand that BBMP 

> Launch a Trade License collection drive in all major market and commercial areas of Bangalore.


## Hotels

* Travel portals  list about 1070 Hotels in Bangalore. This does not include the smaller boarding and lodging units. 
* BBMP's Trade license data shows that Bangalore has
   + 649 lodging units. i.e. _atleast `r 100 - 100*649/1070`% units are missing from BBMP trade license dataset_
   + 783 PG Units i.e. 4 PG Homes per ward. _Visual inspection and the scourge of PG Home pamphlets on every street indicates a different picture. Check below for a detailed analysis of PG Houses_
   + 306 Choultries, Marriage halls, Party halls, convection centers etc across Bangalore. i.e. approx 1.5 per ward. _With every other restaurant sporting a party hall and a random google search throwing up series of kalayan mantapa's within less than a Km of each other and religious institution also doubling up as marriage halls the actual number will be far higher_

##  We demand that BBMP 

* Compare the Hotel data from the travel sites to its internal license data and issue demand notice to those units operating without proper trade license.
* Extract the location of kalyan mantapa's from the marriage registration database and issue demand notice.
* Issue demand notice to all restaurants sporting a party hall.


## Paying Guest accomodations in Bangalore
```{r}
pgHomes<- df_trade_license %>% group_by(Ward.Name,WARD_NO)%>%filter(grepl("Paying Guests",Trade.Sub.Name)) %>% summarise(NumPG=n(),fees=sum(Amount.Paid))%>%arrange(NumPG)
```
BBMP's Trade license database reports `r sum(pgHomes$NumPG)` Paying Guest homes in `r NROW(pgHomes)` Wards. BBMP has either not collected the license fee or is not reported the collection in the remaining `r 198 - NROW(pgHomes)`



```{r echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
shp<-readOGR("../Data-Driven-Policy/book/bangalore/bangalore/gis/bbmpwards","bbmpwards")
shp@data<-left_join(shp@data,pgHomes)
shp@data<-shp@data%>%mutate(popup=paste0("Ward=",WARD_NAME,"<br/>PG=",NumPG))
```
```{r}
pal<-colorBin(rev(topo.colors(10)),shp$NumPG,na.color="#FF0CC0")
leaflet(shp)%>%
addTiles(urlTemplate = "http://{s}.basemaps.cartocdn.com/light_all/{z}/{x}/{y}.png") %>%
addPolygons(stroke=T,fillColor =pal(shp$NumPG), popup=shp$popup)  %>%
#addCircles(lng=shp$LON,lat=shp$LAT,radius=shp$NumPG)%>%
addLegend(position = 'topright', pal=pal,values=shp$NumPG, labels = "PG Homes across wards", opacity = 0.4,       title = 'Legend')
```

* Examining the map in detail we find examples like 
  + Kumarswamy layout with 20 PG Homes, with its neighboring wards having 0-3 PG Homes.
  + Madiwala with 88 PG Homes while it neighbor Jakkasandra has only 4 and Gurappanapalya has 0 PG Homes.

If we assume that proper identification of PG homes took place in wards with more than 10PG Homes then the average across Bangalore would be `r mean(pgHomes[pgHomes$NumPG>=10,]$NumPG)` instead of the current average of `r mean(pgHomes$NumPG)` PG Homes.

```{r results='asis'}
 print(xtable(pgHomes),type="html")
```


## Where is the Building Mafia and the Water Tanker Mafia?

* BBMP Has a category of Construction materials(brick, sand, cement, roofing etc. ) under which only 81 Traders are registered.
* Similarly there is only one registered Water Tanker operator in Bangalore.

## Other gross Errors...

* 13 Firework wholesale/retail units are listed. Every Diwali we have at least 4-5 fire cracker shops in our layout averaging to approximately 20 shops across the ward. Extrapolating this, we should have around 4000 shops across Bangalore
* BBMP lists 6 theaters whereas Bookmyshow lists 93 theaters.

# Conclusion

## Revenue Leakage

We have identified 6 types of revenue leakage in BBMP's System

1. Units who do not pay.
2. Units which are mis-classified and pay at lower rate.
3. Units who pay but are issued fake challan.
4. Reuse of the same DD number for multiple units in the system
5. Non-remittance of the amount collected to the proper account.
6. Delayed remittance of the amount collected to the proper account.
7. Non reversal of Bounced cheque transaction in the system.
8. Manipulation of due amount in the system backend.

The first two require a collusion between the citizen and the BBMP revenue officer. The collusion may require something as simple as BBMP officers neglecting their duties for this type of leakage to take place. or it may be as worse as the BBMP officers being active participants and advisors to the citizens in this evasion scheme.

The third, fourth, fifth and sixth can be addressed by eliminating the need for the citizen to interact with the BBMP officials. Direct online transfer can address this issue. Merging the old records in the new system and enabling online payment of past dues is also required to curb this set of leakage.

The seventh type of leakage can be detected by regular reconciliation of BBMP accounts with the corresponding bank.

The eights type of leakage can be addressed by auditing the tax/fees computation and tax/fees collection systems.

## We Demand that BBMP 
> Should publish a white paper with detailed timebound  remedial actions that will be taken for each of the leakage.


The amount collected per month is
```{r results='asis'}
df_amount<-df_trade_license%>%group_by(month)%>%summarise(Amount=sum(Amount.Paid))
ggplot(df_amount,aes(x=month,y=Amount))+geom_line()
```

* In its [Press Release](http://www.newindianexpress.com/states/karnataka/Steep-Hike-in-BBMP-Trade-Licence-Fee-From-Feb-1/2016/01/29/article3250355.ece) BBMP claims that it has collected 18 Crores from 55,000 Traders.
* Data Available indicates they have collected `r sum(df_amount$Amount)/crore` Crore from `r NROW(unique(df_trade_license$application..Number))` traders. We request BBMP to put the remaining data online.
* The analysis of available Data shows that the reported fees collection is far less than the potential.
* The average fees collected is `r mean (df_trade_license$Amount.Paid)` targeting the missing 4 lakh traders would result in a total collection of at least `r mean(df_trade_license$Amount.Paid)*455000/crore` Crores.
* This year BBMP has increased the fee rates (at least doubled for all categories, 4 times increase for a few category) So the potential collection for 2016-2017 will be `r 2*mean(df_trade_license$Amount.Paid)*455000/crore` Crores


# Appendix

### Property Tax Residential Units

* The city population is currently estimated to be over 1 crore.
  + This translates to over 25 lakh residential units (average family size of 4 as per census)
* As per BBMP there are over 16 lakh residential properties in Bangalore.
* Last year BBMP Collected property Tax from 10 Lakh residential properties.
  + BBMP's revenue loss from residential properties is between 40-60%
* As per our sources while over 2 lakh residential units were built from 2010 to 2015 only 95 Occupancy certificates were issues.

## Demand 1

> **Demand** In accordance with Section 4-1b xiii, We demand that BBMP proactively and regularly declare 
> 1. The details of the permits and authorizations granted by it for construction of the residential and commercial units in Bangalore,
> 2. The details of the authorization to occupy a residential unit.




### Commercial Units.

# Trade license

## Parting Shot

## Ayurvedic clinics or Barber shops?
BBMP classifies Ayurvedic clinics as either Barber Shops or grocery/food stores. 

* Does it mean that these clinics are not authorised to treat patients?
* Or is BBMP unable to distinguish between ancient Indian Medical system and Ancient European medical system?
```{r results='asis'}
clinics<-df_trade_license %>%  filter(grepl("Ayurv",Trade.Name))%>%
                select(Trade.Name,Trade.Minor.Name)
 print(xtable( clinics   ),type="html")

```


## The heads under which fees are collected are.

```{r results='asis'}
 print(xtable(table(df_trade_license$Trade.Minor.Name)),type="html")

```


# Missing Items
The following heads have no receipts..

* Canteen in industries/hospitals etc. 
* Heads for building material (lime,cement, sand, stone etc.) 

### 3rd Party sources
#### Make My Trip
* Make my trip reports 1070 hotels in Bangalore
* 39 5 star hotels
* 57 4 star hotels
* 261 3 star hotels
* 121 2 star hotels
* 300 1 star hotels

#### Zomato
* Zomato lists 8000+ restaurants in Bangalore

#### JustDial 
* JustDial has 319 listings under Kalyan Mantapa. This will not cover the smaller party halls attached to restaurants.

# Explore this dataset.
The script and dataset used to create this analysis are available at
https://github.com/DataDrivenPolicy/BBMP-Finances
